### 2017-09-10 writeup/cover_letter.tex
added cover letter

\documentclass[11pt]{letter}
\usepackage[hidelinks]{hyperref}
\usepackage[green]{msuletterhead}
%\topmarginskip{-0.35in}
%\bottommarginskip{-1.5in}
%\leftmarginsize{1in}
%\rightmarginsize{1.25in}
%\sigskipbefore{0.4in}
%\sigskipafter{0in}
%\noLines
%\nolines
%\noHeadline
%\noheadline
%\signature{Gideon Bradburd, Graham Coop, and Peter Ralph}
%
%\namefrom{}
%\addrfrom{Gideon Bradburd \\ Dept. of Integrative Biology \\ Michigan State University \\ East Lansing, MI 48823 USA \\ 
%\href{mailto:bradburd@msu.edu}{bradburd@msu.edu}}
%\emailfrom{bradburd@msu.edu}
%
%\greetto{To the Editor(s) --}
%\closeline{Sincerely,}

\begin{document}
%\begin{newlfm}

Submission Cover Letter\\
\emph{PLoS Biology}
\\\\\\
To the Editor(s),

Please find enclosed our manuscript,  
``Inferring Continuous and Discrete Population Genetic Structure Across Space,"
for consideration for publication in PLoS Biology.

In the manuscript, we present a novel statistical method 
for the simultaneous inference of continuous and discrete patterns of population structure.
The method estimates ancestry proportions for each 
sample from a set of discrete population layers, 
and, within each layer, estimates a rate at which relatedness decays with distance.
This model explicitly addresses the ``clines vs. clusters" problem in 
modeling population genetic variation by jointly accommodating 
continuous and discrete patterns of differentiation. 
We demonstrate the utility of this approach using both simulations 
and empirical applications, with datasets from poplars and black bears. 
We also release an associated R package, called \texttt{conStruct}, 
for ``\emph{con}tinuous \emph{struct}ure."

This work is motivated by two observations:
first, that patterns of isolation by distance (IBD) are ubiquitous in nature; 
and second, that existing model-based clustering methods, 
such as \texttt{STRUCTURE} (Pritchard \emph{et al.} 2000) 
or \texttt{ADMIXTURE} (Alexander \emph{et al.} 2009), 
can only describe continuous patterns of variation using discrete clusters, 
and so must use a possibly large number of clusters
to describe continuous population differentiation.
Inference of these spurious clusters can throw a 
wrench in the gears of efforts to describe and categorize natural variation, 
one of the first steps in any analysis of genetic data, 
and a principal mission of biology.

In this manuscript,
we use simulations and empirical data to demonstrate 
the ways in which nonspatial clustering approaches 
fail in the presence of IBD; 
we also present two different metrics for choosing an appropriate 
number of layers to describe the data, 
and, using both, 
are able to establish that our method performs quite well.

As of this submission, \texttt{STRUCTURE} alone has been cited more than 20,000 times, 
and the falling costs of sequencing are facilitating an explosion of next-generation sequence datasets.  
The first step in the analysis of these data will, for most researchers, 
involve a model-based clustering analysis.
We therefore believe that our work represents an important and timely contribution to the field.

John Novembre (jnovembre@uchicago.edu), \\
Desislava Petkova (desislavka@gmail.com), \\
Jonathan Pritchard (pritch@stanford.edu),  \\
Sohini Ramachandran (sramachandran@brown.edu), \\
and Matthew Stephens (mstephens@uchicago.edu) \\
would be suitable reviewers.

Sincerely,
Gideon Bradburd, 
Graham Coop, and 
Peter Ralph
%\end{newlfm}
\end{document}

%What is the scientific question you are addressing?
%What is the key finding that answers this question?
%What is the nature of the evidence you provide in support of your conclusion?
%What are the three most recently published articles that are relevant to this question?
%What significance do your results have for the field?
%What significance do your results have for the broader community (of biologists and/or the public)?
%What other novel findings do you present?
%Is there additional information that we should take into account?