### 2018-05-21 writeup/response2reviews.tex
started edits for genetics reviews

\documentclass[11pt]{letter}
\usepackage[hidelinks]{hyperref}
\usepackage[green]{msuletterhead}
\usepackage[usenames,dvipsnames]{xcolor}
\pagenumbering{gobble}

\newcommand{\gb}[1]{{\bf\color{black}{#1}}}
\newcommand{\plr}[1]{{\it\color{purple}{(#1)}}}
\newcommand{\gc}[1]{{\it\color{blue}{(#1)}}}

\begin{document}

Re-Submission Cover Letter \hfill	\today\\
\emph{Genetics}
\\\\\\
To the Editor,

Please find enclosed the re-submission of our manuscript, 
``Inferring Continuous and Discrete Population Genetic Structure Across Space,"
for re-consideration for publication in Genetics.

We appreciate the opportunity to submit a revised version of the manuscript.  
The comments provided by the Editor 
and the anonymous reviewer are very helpful, 
and we believe the manuscript is improved as a result 
of their critiques.

Our detailed responses to those comments are inset below (in bold), 
but, in brief, we have included more of the results of the ADMIXTURE analyses 
in the main body of the text.
Specifically, we have added figure panels to Figures 7 and 8 
that explicitly compare \texttt{conStruct}'s results with those of ADMIXTURE 
on the black bear dataset.
We have also added text throughout the results to increase the depth of 
our comparison of \texttt{conStruct} and ADMIXTURE.

We hope that the paper is now acceptable for publication.\\\\
%

Thank you for your consideration,\\
Gideon Bradburd (for all of the authors).

\newpage

\textbf{Editor}\\
We received the previous reviews for your paper, and sent also sent it out to a new reviewer, along with your response to the earlier reviews. Both the new reviewer and I believe that you have answered the previous criticisms in a reasonable way, and I am pleased to tell you that your paper is potentially suitable for publication in GENETICS. However, the new reviewer did have several comments which you should consider in preparing the final version (see the end of this email). In particular, they suggest including more of the ADMIXTURE models in the main text. I think that this could be helpful, but would not insist on it.

\textbf{Reviewer}\\
\begin{itemize}
\item This manuscript presents a valuable model that addresses an long standing chestnut in population genetic analysis, which is how to jointly analyze discrete and spatially continuous variation within a single dataset.  The method presented does seem to be something of a breakthrough in addressing this problem sensibly.  I found the authors response to previous reviewer comments to be generally reasonable. However, I do think that there is a strong case for putting some more ADMIXTURE model results into the text and possibly for some additional analyses.

\item I agree with the authors point in their letter that ADMIXTURE/STRUCTURE is important because it is what almost everyone would do otherwise.  The key goal for the manuscript is to show that the method that everyone uses gives results that can be misleading/unhelpful and that the layered model is a substantial improvement. This is very clear for the simulated data. The previous reviewers are right to ask for additional simulations in the sense that a more biological simulation would make a stronger case for practical utility and it need not lengthen the manuscript significantly, since the current simulations could be largely relegated to supplementary material as representing "simulation according to the model". But this is definitely an optional, since there are presentational arguments to be made either way.

\item A more biological simulation is not necessary as long as the real data examples are convincing. For the Poplar data, the key insight of the model is that there are three well-supported layers in addition to isolation by distance. This is not clear according to ADMIXTURE, which gives continuous improvement for K, although the clustering itself is reasonable. The differences in patterns of IBD between populations are also interesting.

\item For the bear data, there is a stronger case to be made, namely that as well as estimating a biologically insightful number of layers, individual ADMIXTURE or non-spatial model components are positively misleading for K=3 or more. For the Poplar example, the results are not actively misleading , they just represent finer and finer geography. The misleadingness of the bear results can be shown by the fact that although the cross validation shows continuous improvement, the admixture components are highly inconsistent between runs. For example at K=3, a component is shown that is shared between East and West, which if interpreted literally would represent a biologically important result about connections between populations but the clustering of the yellow populations disappears entirely for K=4. Therefore, I think there is a good case to be made for showing the ADMIXTURE model results in the text and to discuss the misleading/ unstable nature of any particular result.

\item It would be interesting to do some kind of plot for ADMIXTURE results, to demonstrate that populations that share affinity according higher components are not particularly similar when allowing for geographical distance. This would be a rather direct way of showing that the ADMIXTURE model is not producing insightful results, despite the continuous improvement in model probability for higher K. In this context, I always think of the "clines vs clusters" paper by Rosenberg. It would be great to have some kind of figure to show that for K=2, it looks like clusters but for K=3, the clusters add much less.

\item The layer specific covariances show a very funny pattern, tending to infinity for low distances for components 3 and above for the bear data. I am not sure I understand this. Is this discussed anywhere?\\\\
\gb{I think what's happening is that these layer-specific covariances for layers $>$2 
are describing local departures from global patterns of IBD, 
as might be expected due to, e.g., inbreeding.
We have added some text to that effect in Lines XXX-XXX.}

\item Minor point, in figure 4, legend, should it read in each run of the spatial model, rather than nonspatial model?\\\\
\gb{Good catch!  Caption has been fixed.}
\end{itemize}

\end{document}

\begin{itemize}
\item First, many results concern comparisons between clustering based on the spatial and non-spatial models of construct (Figs 2, 3, 4, 6, 7, 9). However the authors should rather compare their spatial approach to existing software for model-based clustering, which do not account for space (Structure, fastSTRUCTURE, ADMIXTURE,...). Results obtained with one of these software should not be available in supplementary material but should rather be visible in the main text.\\\\
\gb{The utility of the nonspatial model implemented in \texttt{conStruct} 
is that its performance can be directly compared to that of the spatial model; 
their cross-validation predictive accuracy values are in the same currency 
because they use the same Wishart likelihood model.
So, using it we can say not just that that the nonspatial model prefers 
$K=X$ while the spatial model prefers $K=Y$, 
but also, e.g., that the spatial model is preferred over the nonspatial model.
However, as you note, it's important to compare \texttt{conStruct}'s performance to 
that of existing nonspatial model-based clustering methods as well, 
which is why we had also included analyses using fastSTRUCTURE.

\ Your point here is well-taken though, because the fastSTRUCTURE analyses, 
which included no model comparisons, did not provide a direct basis of comparison to \texttt{conStruct}.
There is a model comparison tool implemented in fastSTRUCTURE, 
but we were unable to get it to run on our datasets.
We have contacted the fastSTRUCTURE authors 
and alerted them to this issue, 
but have yet to hear back from them.

\ We have therefore analyzed all datasets included in the paper using ADMIXTURE, 
and used ADMIXTURE's cross-validation procedure to compare model performance across values of $K$.  
These results are discussed extensively throughout the updated manuscript ``Results" section; 
and the model that minimizes the cross-validation error in the ADMIXTURE analyses on simulated data 
is highlighted in the the simulation cross-validation results figure.
However, due to the already extensive length of the manuscript (which Reviewer 2 points out), 
the results of the ADMIXTURE analyses are given in the Supplementary Materials.
}\\

\item Second, the simulation scenario should correspond to a real biological scenario. Here this layer-based simulation scenario is very difficult to understand. I would be more interesting to stick to more classic spatially explicit simulations such as range expansion, or secondary contact scenarios.\\\\
\gb{We appreciate that the simulation scenario we ran is somewhat artificial, 
and that it would be good to test the behavior of the method 
under a classic ``isolation followed by expansion into a zone of secondary contact" model.
However, as we discuss in the manuscript in the section ``Simulation details",
we wanted to simulate datasets that had unambiguous true admixture proportions, 
so that we could test \texttt{conStruct}'s ability to estimate admixture accurately 
(shown in SuppMat figures 8 and 13).
A more organic simulation, with isolation followed by expansion, 
secondary contact, and subsequent admixture, 
would not facilitate that test in a straightforward way. 
As mentioned by Reviewer 2, there are an enormous number 
of possible simulations we could run to test \texttt{conStruct}'s performance, 
but fully exploring them is beyond of the scope of this manuscript, 
so we intend to pursue those in future work.}\\

\item Third, I think that the main point of the paper should be more explicit. Non-spatial model based approaches are prone to overfitting when K is too large and this can be avoided when using spatial-based approaches. \\\\
\gb{
Good suggestion; we've added a line to the Abstract to this effect.
}\\

\item Last, there are other software that account for space when performing model-based clustering and they should be included in the comparisons (BAPS, Geneland, TESS,...). Right now, they are only mentioned in the introduction and we would like to see results comparing these different model-based approaches.\\\\
\gb{
The methods mentioned here --- BAPS, Geneland, and TESS --- 
place spatial priors on admixture proportions
and/or cluster membership, 
but like STRUCTURE, they assume that allele frequencies 
within a cluster are constant across space.
Therefore, although spatially explicit, this parameterization does not 
account for space in a way that avoids the problem we describe in our manuscript: 
the characterization of spatially continuous differentiation 
as clines of admixture across spurious clusters.

\ A comprehensive comparison between \texttt{conStruct} 
and all other model-based clustering methods is beyond the scope of this paper. 
As STRUCTURE is the canonical model-based clustering algorithm
(having been cited $>$ 20,000 times), but is too slow to be run 
on the modern-scale SNP datasets we simulate, 
we feel that ADMIXTURE, which implements an extremely similar model 
within a faster inference framework, 
provides the best comparison to our method.}\\

\item In the introduction, the authors write ``placing spatial priors on cluster memberships'' when referring to other methods that deal with ancestry proportions instead of cluster memberships.\\\\
\gb{
Guillot \emph{et al.}, in the 2005 paper that introduced the method Geneland, 
refer to this quantity as ``population membership."
We kept the wording consistent for clarity, but changed ``population" 
to ``cluster" to keep a distinction between biological and statistical groupings.}\\

\item Formula (1) about the covariance is difficult to understand. It does not correspond to a classic definition of a covariance. Does this correspond to the covariance of some random variables? The discussion following the definition of the covariance looks like a discussion and should be moved to the discussion.\\\\

\gb{The definition given is a covariance of some random variables; we have made this fact more clear, 
and we have substantially rewritten the section on the allelic covariance
and have moved some of the intuition-building discussion to the Materials and Methods.
We have also moved some of this discussion to the Discussion (heading ``Allelic or genetic covariance?'').
}\\

\end{itemize}

\textbf{Reviewer 2}\\
In this paper, Bradburd et al. present a novel analysis method for the joint characterization of discrete and continuous structure. Their model uses a mixed membership Gaussian process, adding a within layer (cluster) Isolation-by-distance-process. I do think that this paper is very clearly written, and presents the model and method in great detail. Besides the presentation of the method, a simulation and two application studies are performed, that lead to a comprehensive overview of the method. I would consider this model to be very valuable for a large audience in ecology and evolution, including biogeographers and conservation biologists, with the only drawback being the very high computational requirements. Overall, however I recommend the paper for acceptance with only very minor changes.

\begin{itemize}
\item One issue I am wondering about is whether there are some issues with ``outlier individuals'', that e.g. represent poorly sampled clusters, which might be an issue with the Alaskan/NW-Canadian bears. In particular, consider the case where the two layers were only fitted from the samples from the 48-states. Presumably, the resulting covariance curve would be similar to Fig S26b (assuming the two layer case). If we then added those NW-pacific samples, we might expect them to be most likely added to the cluster with the flatter right tail, even if they were essentially unrelated. This issue can also arise in kriging (Figure 6 in the first version of the preprint of Caye et al (on bioRxiv). Could this interpretation be excluded somehow?
\\\\
\gb{This is a great point, and highlights a subtle problem.
Although we feel that isolation-by-distance is a reasonable model for population differentiation, 
it is possible that, by implementing a spatial model, 
we can facilitate ``spurious" layer membership in far-flung samples, 
who are sampled far enough away from all other samples 
that their spatial covariance is negligible.  
This situation is made less likely by the $\phi$ parameter in each layer, 
which induces some amount of covariance between samples in a layer 
regardless of their geographic separation.
However, it is possible that $\phi$ in a layer might be estimated to be $\approx 0$. 
We've discussed heuristics for highlighting such samples 
(e.g., highlighting pairs of samples with $>$50\% membership in a given layer,
 that, given their geographic separation, 
 the parametric covariance curve in that layer, 
 and the $\phi$ parameter in that layer, have effectively zero covariance), 
 but we have not yet implemented or tested these checks.
We are also currently developing a set of spatial \textit{f}-statistics, 
analogous to those presented in Reich \emph{et al.} (2009) and discussed further in Peter (2016), 
which could be useful in highlighting such ``outlier" samples.
But, as you note, this manuscript is already quite long, 
so this work will be a part of future projects.
}\\

\item Another thing that could be added are some additional simulations in more complex demographic models, such as expansions, temporally/spatially heterogeneous migration, or sparser sampling schemes. However, as the paper is already rather long (at almost 20 single-spaced pages), and the space of scenarios for which the model is relevant is enormous, I hope the authors consider further exploring their model in a follow-up study.\\\\
\gb{We completely agree, and further explorations of the behavior of the method 
in different scenarios are being planned now.}\\

\item 100: Is there a reason that allelic covariance is used instead of genetic covariance?\\\\
\gb{
    Good question; we've added some more text for this, 
    both explaining what the allelic covariance is and why we use it in lieu of the genetic covariance,
    (in the ``Data" subsection of the ``Results" 
    and the ``Allelic covariance" subsection of ``Materials and Methods").
    The short answer is that the allelic covariance is insensitive to the choice of reference allele, 
    and, because it doesn't subtract off means like the genetic covariance, 
    has a constant value between a pair of samples in the dataset, 
    regardless of which other samples are included.
}\\

\item Could supp. figures be sorted according to first mentioning in ms?\\\\
\gb{Fixed.}\\

\item Fig3: increase label size\\\\
\gb{Fixed.}\\

\item Last line in paragraph on p. 18: as the derivations are otherwise very detailed and easy to follow, it might be worth adding here that this follows from E[UV] = 4 Cov[A, B]\\\\
\gb{
We have substantially rewritten this section for greater clarity.\\
}

\item 634: Using ms,
\\\\
\gb{Fixed.}\\

\item Fig S4: Could the colors be made to correspond each other (e.g. panel e vs 3). In general, is the exchangeability of layers a practical concern, or similar to the original structure, does the MCMC never ``switch'' two layer labels?\\\\
\gb{In STAN, we declare the vector of $\phi$ parameters to be an object of class ``positive ordered",
	meaning that values of phi must all have values greater than zero, 
	and must be in strictly increasing order.
	In practice, I don't think that much label-switching would occur within the course of a single MCMC run 
	(like in STRUCTURE), but inducing a specific order to the $\phi$ parameter makes that less likely still.
	There can be label switching across independent runs for the same value of $K$, 
	and of course there's label switching across runs with different numbers of $K$.
	In the \texttt{conStruct} R package, we've included a function that uses the cluster matching 
	algorithm described in CLUMPP (Jakobsson and Rosenberg 2007) to match layers across runs.
	We used that algorithm to keep plotting colors consistent in our figures, including in Fig. S4.
	However, something funny happens in that sequence of analyses between $K=5$ and $K=6$, 
	where the ranking of cluster matching algorithm gives wonky results, 
	leading to the non-intuitive layer color matching.
	We've tweaked that one by hand to keep the plotting nice.
 }\\

\item Fig S26: maybe keep Y-axis scale consistent between panels?\\\\
\gb{Fixed.}\\

\item Cross-Validation: As we all know SNP are not uncorrelated. I am wondering if that could influence the CV, as spurious patterns in the testing data might be partially explained by nearby SNP in the training data. Is that a concern?\\\\
\gb{
	This is a good point and it definitely could be a concern.
	We've included a discussion of this in the cross-validation procedure description, 
	as well as a description of the CV procedure that should be used if you know LD between loci.
	But, we also imagine that, for a while at least, the vast majority of \texttt{conStruct} users 
	won't have genomic position information for their SNP data.
}\\

\end{itemize}

\end{document}